Fixing Vulnerabilities Hinders Maintainability

### 2020-01-30 paper/icpc20.tex
new rq

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

@@

  \url{https://figshare.com/s/3e17518222df6a11a8d6}.

@@

\medskip
\begin{lstlisting}[style={CStyle}, caption={Fix provided by OpenSSL developers to the
\\CVE-$2014$-$3506$ vulnerability},label={lst:vuln}]
+ static unsigned long dtls1_max_handshake_message_len(const SSL *s){ /*!\annotation{lst:func1}!*/
+  unsigned long max_len = DTLS1_HM_HEADER_LENGTH + SSL3_RT_MAX_ENCRYPTED_LENGTH;
+  if(max_len < (unsigned long)s->max_cert_list)
+     return s->max_cert_list;
+  return max_len;
+ }

@@

\emph{Better Code Hub} tool. The same approach is applied to a randomly generated
dataset of regular commits --- baseline --- that we use to understand how
they are not.

@@

\textit{\textbf{RQ2} Which patterns of security patches are more likely to

@@

\begin{figure}[h]	
 	\centering 	\includegraphics[width=0.46\textwidth]{figures/maintainability_sec_vs_reg.pdf}
 	\caption{Maintainability difference for security and baseline patches.}
	\label{fig:secvsreg}

@@

\section{Threats to Validity}\label{sec:threats}
%
The following section presents the potential threats to the validity of this
study.
%
\subsection{Construct}
%
The maintainability formula was inferred based on the BCH's reports. The high
amount of different projects and backgrounds may require other
maintainability standards. However, BCH does use a representative benchmark of
closed and open-source software projects to compute the
thresholds for each maintainability guideline~\cite{Visser:2016:OREILLY, Baggen2012}.

\subsection{Internal}

The security refactorings dataset provided by the previous
work~\cite{Reis:2017:IJSSE} was collected based on the messages of GitHub
commits produced by project developers to classify the changes performed by them
to fix security flaws. This approach discards security refactorings that were
not explicit in commits messages. Baseline commits are retrieved randomly by selecting one commit from the same
project of a given security refactoring. This approach softens the differences
that may result from the characteristics of each project. However,
maintainability may still be affected by the developers' experience, coding
style and software contribution policies which is not evaluated in this study.
Furthermore, this evaluation considers that $607$ regular commits are enough to
alleviate random irregularities in the maintainability differences of the
baseline.

\subsection{External}

We use a dataset that only comprises refactorings of open-source software.
However, the methodology requires to access data that is not publicly available.
Thus, our findings may not extend to non-open source software. Different programming 
languages may require different coding practices to address software safety. The 
dataset comprises more commits in C and PHP, i.e., the dataset may not be representative 
of the population regarding programming languages. Our approach does not consider 
security refactorings in any other languages but English.

>>>>

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
%\begin{CCSXML}
%<ccs2012>
% <concept>
%  <concept_id>10010520.10010553.10010562</concept_id>
%  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
%  <concept_significance>500</concept_significance>
% </concept>
% <concept>
%  <concept_id>10010520.10010575.10010755</concept_id>
%  <concept_desc>Computer systems organization~Redundancy</concept_desc>
%  <concept_significance>300</concept_significance>
% </concept>
% <concept>
%  <concept_id>10010520.10010553.10010554</concept_id>
%  <concept_desc>Computer systems organization~Robotics</concept_desc>
%  <concept_significance>100</concept_significance>
% </concept>
% <concept>
%  <concept_id>10003033.10003083.10003095</concept_id>
%  <concept_desc>Networks~Network reliability</concept_desc>
%  <concept_significance>100</concept_significance>
% </concept>
%</ccs2012>
%\end{CCSXML}
%
%\ccsdesc[500]{Computer systems organization~Embedded systems}
%\ccsdesc[300]{Computer systems organization~Redundancy}
%\ccsdesc{Computer systems organization~Robotics}
%\ccsdesc[100]{Networks~Network reliability}

@@

  \textit{<anonymized for double blind review>}.

@@

\medskip
\setcounter{lstannotation}{0}
\begin{lstlisting}[style={CStyle}, caption={Fix provided by OpenSSL developers to the
\\CVE-$2014$-$3506$ vulnerability},label={lst:vuln}]
+ static unsigned long dtls1_max_handshake_message_len(const SSL *s){ /*!\annotation{lst:func1}!*/
+  unsigned long max_len = DTLS1_HM_HEADER_LENGTH + SSL3_RT_MAX_ENCRYPTED_LENGTH;
+  if(max_len < (unsigned long)s->max_cert_list)
+     return s->max_cert_list;
+  return max_len;
+ }

@@

\emph{Better Code Hub} tool. 

@@

\textit{\textbf{RQ2} What is the impact of security patches versus regular changes on the
We also compute the maintainability of $1037$ random regular commits using the
\emph{Better Code Hub} tool --- baseline.  We use them to understand \textbf{how
they are not}.
\textit{\textbf{RQ3} Which patterns of security patches are more likely to

@@

\begin{figure}[h]	
 	\centering 	\includegraphics[width=0.43\textwidth]{figures/maintainability_sec_vs_reg.pdf}
 	\caption{Maintainability difference for security and baseline patches.}
	\label{fig:secvsreg}

@@

\section{Threats to Validity}\label{sec:threats}
%
The following section presents the potential threats to the validity of this
study.
%
\\\textit{Construct:} The maintainability formula was inferred based on the BCH's reports. The high
amount of different projects and backgrounds may require other
maintainability standards. However, BCH does use a representative benchmark of
closed and open-source software projects to compute the
thresholds for each maintainability guideline~\cite{Visser:2016:OREILLY, Baggen2012}.
%
\\\textit{Internal:} The security refactorings dataset provided by the previous
work~\cite{Reis:2017:IJSSE} was collected based on the messages of GitHub
commits produced by project developers to classify the changes performed by them
to fix security flaws. This approach discards security refactorings that were
not explicit in commits messages. Baseline commits are retrieved randomly by selecting one commit from the same
project of a given security refactoring. This approach softens the differences
that may result from the characteristics of each project. However,
maintainability may still be affected by the developers' experience, coding
style and software contribution policies which is not evaluated in this study.
Furthermore, this evaluation considers that $607$ regular commits are enough to
alleviate random irregularities in the maintainability differences of the
baseline.
%
\\\textit{External:} We use a dataset that only comprises refactorings of open-source software.
However, the methodology requires to access data that is not publicly available.
Thus, our findings may not extend to non-open source software. Different programming 
languages may require different coding practices to address software safety. The 
dataset comprises more commits in C and PHP, i.e., the dataset may not be representative 
of the population regarding programming languages. Our approach does not consider 
security refactorings in any other languages but English.
